### 2024-02-09 outputs/Paper/paper.qmd
generated graphs

---
title: "Do Economics Majors Make More Than Other Majors?"
author: Diana Liu
thanks: "Code and data are available at: https://github.com/Diana-Guanzhi-Liu/Do-Economics-Students-Make-More-Than-Other-Majors-"
date: 11 February 2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(gt)
library(shiny)
library(ggplot2)
library(scales)
library(gridExtra)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....
(@CitePaper)

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: tbl-constructed data
#| fig-cap: Sample of Constructed data from Bleemer and Mehta
#| echo: false
load("/cloud/project/Inputs/Raw Data/ACS_Data_UCSCEcon_NAICS.Rda")
gt(head(acs))

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: tbl-cleaned data
#| fig-cap: Sample of Cleaned Data of UCSC Students
#| echo: false
#| warning: false
#| message: false

load("/cloud/project/outputs/Data/Majors_Income_Cleaned.Rda")
gt(head(cleaned_majors_data))

```

Talk way more about it. 
```{r}
#| echo: false
#| warning: false
#| message: false

load("/cloud/project/outputs/Data/Majors_Income_Cleaned.Rda")

#Sort Majors into Categories
Sciences <- c("Agriculture", "Environment and Natural Resources", "Communication Technologies", "Family and Consumer Sciences", "Library Science", "Biology and Life Sciences", "Mathematics and Statistics", "Physical Sciences", "Medical and Health Sciences and Services")
Arts <- c("Architecture", "Communications", "Area, Ethnic, and Civilization Studies", "Education Administration and Teaching", "Cosmetology Services and Culinary Arts", "Linguistics and Foreign Languages", "Law", "English Language, Literature, and Composition", "Liberal Arts and Humanities", "Physical Fitness, Parks, Recreation, and Leisure", "Philosophy and Religious Studies", "Theology and Religious Vocations", "Psychology", "Criminal Justice and Fire Protection", "Public Affairs, Policy, and Social Work", "Social Sciences", "Precision Production and Industrial Arts", "Fine Arts", "History")
Business <- c("Business")
Applied_Science <- c("Computer and Information Sciences", "Engineering", "Engineering Technologies", "Military Technologies", "Nuclear, Industrial Radiology, and Biological Technologies", "Construction Services", "Electrical and Mechanic Repairs and Technologies", "Transportation Sciences and Technologies")

business <- subset(cleaned_majors_data, Label %in% Business)
sciences <- subset(cleaned_majors_data, Label %in% Sciences)
arts <- subset(cleaned_majors_data, Label %in% Arts)
applied_science <- subset(cleaned_majors_data, Label %in% Applied_Science)

```

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| label: fig-majors income
#| tbl-cap: Histogram of Business, Arts, Sciences, and Applied Sciences majors wage income
#| warning: false

#summary(business)

business_hist <-
  business |>
  ggplot(aes(x =INCWAGE)) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  xlim(0, 200000) +
  theme_minimal() +
  labs(title = "Wages of Business Majors", x = "Wage Income")

arts_hist <-
  arts |>
  ggplot(aes(x =INCWAGE)) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  xlim(0, 200000) +
  theme_minimal() +
  labs(title = "Wages of Arts Majors", x = "Wage Income")

science_hist <-
  sciences |>
  ggplot(aes(x =INCWAGE)) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  xlim(0, 200000) +
  theme_minimal() +
  labs(title = "Wages of Sciences Majors", x = "Wage Income")

applied_science_hist <-
  applied_science |>
  ggplot(aes(x =INCWAGE)) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  xlim(0, 200000) +
  theme_minimal() +
  labs(title = "Wages of Applied Science Majors", x = "Wage Income")

grid.arrange(arts_hist, business_hist, ncol=2)
grid.arrange(science_hist, applied_science_hist, ncol=2)

```

```{r}

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References